---
title: "Redefinition of RIGs with new Brady IS"
output: html_notebook
author: Maja Kuzman
---

```{r, message=FALSE}
library(GenomicRanges)
library(data.table)
```

Loading the integration data sets and updating the Brady data set:
```{r}
load("is.Robj")
is$Brady <- NULL
brady <- fread("IntegrationSitesInActivatedALLCells_bushman.txt")
brady <- GRanges(brady$V1, IRanges(brady$V2,brady$V2))
is$Brady <- brady
```

Counting number of integration sites per lists:
```{r}
source("2_AdditionalFile1.R")
TableM1
```

Percentage of IS in genes:
```{r}
invitro = mean(unlist(TableM1[3,c("Lucic","Brady")]))
invitro
patients = mean(unlist(TableM1[3,c("Maldarelli","Cohn","Han","Ikeda","Kok","Wagner")]))
patients
```
```{r}
source("3_Supplementary figure 1a.R")
supplementary1a
```
```{r}
source("4_Supplementary figure 1b.R")
plotSaturation
```


Loading genes, updating number of lists:
```{r}
load("GRCh37AllGenes.Robj")
ovlsinGenes <- sapply(is, function(x)(countOverlaps(geni, x, ignore.strand=T)>0))
geni$numberOfLists <- rowSums(ovlsinGenes)

RIGsnumbers <- data.frame(table(geni$numberOfLists),
           cumsum=sapply(0:7, function(i)sum(geni$numberOfLists>=i)),
           expressed=as.numeric(table(genidf[,.(numberOfLists)])),
           cs2=sapply(0:7, function(i)sum(genidf$numberOfLists>=i)),
           expressed2=as.numeric(table(genidf[gene_biotype=="protein_coding",.(numberOfLists)])),
           cs3=sapply(0:7, function(i)sum(genidf[gene_biotype=="protein_coding"]$numberOfLists>=i))

)
colnames(RIGsnumbers) <- c("NrLists","ExactNrGenesOnLists", "MoreThanListAllGenes", "ExactExpressed","MorethanlistExpressed", "ExactProtCod","MoreThanProtCod")

RIGsnumbers
```

Number of protein coding genes in our data set:
```{r}
genidf$RIGSLISTE
genidf[gene_biotype=="protein_coding",.N,RIGSLISTE]
```

Saving list of protein coding RIGs, and list of protein coding genes with NO integration, for Figure 1A:
```{r}
write.table(genidf[(gene_biotype=="protein_coding")&(RIGSLISTE=="2 or more"),"names"], 
            row.names = F, col.names = F,
            quote=F,
            file="RIGs2ormore.txt")
write.table(genidf[(gene_biotype=="protein_coding")&(RIGSLISTE=="0"),"names"], 
            row.names = F, col.names = F,
            quote=F,
            file="RIGs0.txt")

```

Super enhancers and RIGs:
```{r}
genidf[,.N,.(SuperenhancersA,RIGSLISTE)]
```

Expression and RIGs:
```{r}

limits <- quantile(genidf[gene_biotype=="protein_coding"][i.Mean>0, i.Mean], c(0.1, 0.9))
genidf[, ExpressionGroups:=factor(ifelse(i.Mean<=0,"Not expressed", 
                                         ifelse(i.Mean<=limits[1], "Low10%",
                                                ifelse(i.Mean<=limits[2], "Mid","High10%")
                                         )
    ),
    levels=c("Not expressed", "Low10%", "Mid", "High10%")     
    )]
genidf[gene_biotype=="protein_coding",.N,.(ExpressionGroups,RIGSLISTE)][order(ExpressionGroups, RIGSLISTE)]

```

The rest of the figures are plotted in Master.Rmd document.